Switchgrass Phenology and Genotype-by-Environment Mapping

### 2023-12-28 manuscript/revision2/revision2.tex
manuscript: revision 2 author and citation edits

% TODO: Add custom LaTeX header directives here
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Quarto Template for PNAS Submissions},
  pdfauthor={Author One; Author Two; Author Three},
  pdfkeywords={allele-by-environment effect variation, antagonistic
pleiotropy, photoperiod, cumulative growing degree days, genetic
variation},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\title{Quarto Template for PNAS Submissions}

@@

\author[a,c%
,\equalcont%
%
]{Author One}
\author[b%
,\equalcont%
,\correspond%
]{Author Two}
\author[a%
%
%
]{Author Three}

\affil[a]{Harvard University, Department of Government, Cambridge, 2138}
\affil[c]{Harvard University, Department of Statistics, Cambridge, 2138}
\affil[b]{Yale University, Department of Political Science, New
Haven, 3401}

\leadauthor{One, Two, and Three}

@@

\authorcontributions{T.E.J. designed research. D.B.L. contributed plant
material and resources. J.B., D.B.L., and T.E.J. designed and executed
field experiments. A.R.B., P.A.F., F.B.F., D.B.L., R.B.M., F.M.R., Y.W.,
and T.E.J. hosted field experiments. A.H.M. and L.Z. conducted
statistical and computational analyses. The manuscript was written by
A.H.M., L.Z., and T.E.J. with contributions from all authors.}

\authordeclaration{Please declare any competing interests here.}

\equalauthors{\textsuperscript{\equalcont} A.O.(Author One) contributed
equally to this work with A.T. (Author Two) (remove if not applicable).}

\correspondingauthor{\textsuperscript{\correspond}To whom correspondence should be addressed. E-mail: alicem@uchicago.edu}

@@

\begin{abstract}
The timing of vegetative growth and flowering in plants (``phenological
timings'') multiple depend on environmental cues; thus, phenological
traits can have important genotype-by-environment (GxE) interactions.
Here, we map genotype-by-weather effects on phenological timings in
multiple populations of two highly divergent switchgrass (Panicum
virgatum) populations spanning the central United States. We distinguish
GxWeather patterns covarying with interpretable weather-based cues and
agnostic, empirically-inferred GxWeather patterns. Most GxWeather
effects belong to the latter category and do not covary with
weather-based cues. However, \textgreater26\% of Gulf population effects
on flowering time covary with photoperiod cues, while
\textgreater{}\textbf{\% of Midwest upland population SNPs affecting
flowering had effects that covaried with } cues. An independent
pseudo-F2 cross of Gulf and Midwest individuals mapped 23 additive QTLs
for flowering at the same common gardens, all with significant mash
associations and ten with enrichment of highly significant mash
associations. We demonstrate that we can identify QTL with GxE and
assign them to specific weather-based cues and/or data-driven patterns.
Breeding for particular alleles at these loci could change flowering
responsiveness to photoperiod cues in switchgrass. More broadly, this
approach could be used to identify genetic marker-environment
interactions in any species with related populations phenotyped in
multiple environments.
\end{abstract}

@@

\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}\firstpage[20]{2}

@@

Flowering time, in particular, is a common subject of GxE research
(5--11), a key output of selection driving adaptation to local
environments (3, 12, 13), and a selection target for crop improvement to
adapt crops to local or future environments (14). Changing flowering
responsiveness to photoperiod cues has allowed geographic range
expansion and increased yields in a number of cereal species (15--19)
and other crops (20, 21). Recent statistical advances in studying
phenological GxE have involved determining critical environmental
indices before the phenological event occurs, such as photothermal time
within a critical growth window (10). However, most studies of flowering
GxE focus on finding a single, best fitting form of genotype-environment
covariance, despite the key expectation that different genetic
subpopulations, and even different genomic regions, have likely evolved
distinct patterns of GxE. Additionally, despite theoretical predictions
that local adaptation should involve antagonistic pleiotropy, or
sign-changing GxE, at the level of individual loci (22--25), previous
work has found limited evidence of antagonistic pleiotropy (12, 26), but
has been limited by a known statistical bias that reduced detection of
SNP effects that differ in sign (26--28). Thus, despite substantial
interest in the frequencies of various forms of GxE, the prevalence of
antagonistic pleiotropy relative to other forms of GxE remains unknown.

Switchgrass (Panicum virgatum) is considered a short-day plant with
reproductive development strongly linked to day of the year (29).
However, as part of its wide environmental adaptation across the eastern
half of North America, its photoperiodicity has been predicted to differ
by plant latitude of origin (30, 31). We previously found divergent
Midwest and Gulf genetic subpopulations of switchgrass which segregate
for distinct sets of environmental adaptations, based on the analysis of
two measures of fitness (i.e.,biomass and survival) among 732 genotypes
(32). The Midwest genetic subpopulation is primarily composed of
individuals from the well-studied upland switchgrass ecotype (33, 34),
while the Gulf subpopulation has individuals from the lowland ecotype
and the phenotypically intermediate coastal ecotype (32).

Here, we test how the Midwest and Gulf subpopulations differ in their
phenological adaptations across environments, and hence their
phenological GxE. We phenotype a diversity panel of hundreds of
switchgrass genotypes from the Midwest and Gulf subpopulations for the
timing of vegetative development (``green-up'') and reproductive
development (flowering) at eight common garden locations spanning 17
degrees of latitude. These gardens cover the majority of the latitudinal
and climatic range of switchgrass and capture the most comprehensive
picture to date of the environmental variation this species encounters.
After single-site estimation of SNP effects for these phenotypes, we use
multivariate adaptive shrinkage (mash) to jointly estimate SNP effects
in each subpopulation and phenotype at all eight sites (35). Mash allows
us to identify and specify multiple ways genetic effects linked to SNPs
may covary with the environment, and does not have a statistical bias in
detecting SNP effects with the same or opposite signs (39). To confirm
our genetic mapping of GxE, we compare to mapping results from an
outbred pseudo-F2 cross grown at the same sites. Taken together, our
results allow us to describe the environmental cues and genetic
variation affecting phenology in two divergent natural populations of
switchgrass.

@@

In our diversity panel of tetraploid switchgrass (32), genotypes from
the Gulf and Midwest genetic subpopulations had distinct phenological
timings and distinct patterns of phenological correlations across our
eight common garden sites (Fig. 1A,B). At the three Texas common gardens
(hereafter `Texas' gardens), located within the natural range of the
Gulf subpopulation, Gulf green-up occurred before Midwestern green-up,
and Gulf flowering occurred after Midwestern flowering (Fig. 1A). At the
four northernmost common gardens (hereafter `North' gardens), located
within the natural range of the Midwest subpopulation, both Gulf
green-up and flowering occurred after Midwest green-up and flowering. At
the Oklahoma common garden, located near the natural range limits of
both the Gulf and the Midwest subpopulations, Gulf and Midwest green-up
occurred over the same time period. These patterns led to strong
negative phenotypic correlations for green-up between the North and
Texas gardens, particularly in the Gulf and Both subpopulations, and
contributed to positive phenotypic correlations for flowering time of
larger magnitude at more northern gardens (Fig. 1B).

@@

@@

\begin{figure}[H]

{\centering \includegraphics{images/Fig1_200dpi.png}

\caption{Figure 1. Characterization of green-up and flowering dates from
the switchgrass diversity panel. (a) Map and trait histograms of
green-up and flowering dates across two genetically distinct switchgrass
subpopulations and eight common gardens. Purple represents individuals
from the Midwest genetic subpopulation, and pink individuals from the
Gulf subpopulation. Vertical dashed lines indicate the summer solstice.
Common gardens are arranged in latitudinal order. (b) Phenotypic
correlations between clonal replicates planted at eight common gardens,
within and between two genetic subpopulations.~ (c) Narrow sense
heritability of green-up and flowering within single common gardens
(purple) and across all eight common gardens (green), within and between
two genetic subpopulations. (d) Flow diagram of the methods applied to
the green-up and flowering dates to jointly estimate SNP effects across
all sites. Mash was fit to SNP effect data and used to find covariance
matrices that improved the mash model likelihood using a large set of
randomly selected, relatively unlinked SNP effects; this model was
applied to a "strong" set of SNP effects with large effect sizes in the
univariate GWAS.}

\subsection*{Author Affiliations}\label{author-affiliations}
\addcontentsline{toc}{subsection}{Author Affiliations}

Include department, institution, and complete address, with the
ZIP/postal code, for each author. Use lower case letters to match
authors with institutions, as shown in the example. PNAS strongly
encourages authors to supply an \href{https://orcid.org/}{ORCID
identifier} for each author. Individual authors must link their ORCID
account to their PNAS account at
\href{http://www.pnascentral.org/}{www.pnascentral.org}. For proper
authentication, authors must provide their ORCID at submission and are
not permitted to add ORCIDs on proofs.

\subsection*{Manuscript Length}\label{manuscript-length}
\addcontentsline{toc}{subsection}{Manuscript Length}

A standard 6-page article is approximately 4,000 words, 50 references,
and 4 medium-size graphical elements (i.e., figures and tables). The
preferred length of articles remains at 6 pages, but PNAS will allow
articles up to a maximum of 12 pages.

\subsection*{References}\label{references}
\addcontentsline{toc}{subsection}{References}

References should be cited in numerical order as they appear in text;
this will be done automatically via bibtex, e.g.~\cite{belkin2002using}
and
\cite{berard1994embedding,coifman2005geometric,phdthesis,masterthesis}.
All references cited in the main text should be included in the main
manuscript file.

\subsection*{Data Archival}\label{data-archival}
\addcontentsline{toc}{subsection}{Data Archival}

PNAS must be able to archive the data essential to a published article.
Where such archiving is not possible, deposition of data in public
databases, such as GenBank, ArrayExpress, Protein Data Bank, Unidata,
and others outlined in the
\href{https://www.pnas.org/author-center/editorial-and-journal-policies\#materials-and-data-availability}{Information
for Authors}, is acceptable.

\subsection*{Figures and Tables}\label{figures-and-tables-1}
\addcontentsline{toc}{subsection}{Figures and Tables}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3750}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3194}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3056}}@{}}
\caption{Weather variables and time frames around the green up and
flowering date used to construct the hypothesis-based covariance
matrices. The correlations between values of these weather variables for
genetically identical plants grown in different gardens were used to
fill off-diagonal cells of the covariance matrices. Narrow-sense
heritabilities for these values at each garden were used for the
diagonal cells.}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
\textbf{Weather variable}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Time frame}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{\# Covariance Matrices tested in greedy mash algorithm}
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
\textbf{Weather variable}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Time frame}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{\# Covariance Matrices tested in greedy mash algorithm}
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
cumulative GDD at 12C in the time frame & 1, 2, 3, 4, 5, 6, 7, 14, 21,
28 days prior to green up date~ & 60, 30 per phenotype and 10 per
phenotype \& genetic group \\
cumulative rainfall in the time frame & 1, 2, 3, 4, 5, 6, 7, 14, 21, 28
days prior to green up date~ & 60, 30 per phenotype and 10 per phenotype
\&~ genetic group \\
day length (hours) on a specific day indicated by the time frame & 1, 2,
3, 4, 5, 6, 7, and 14 days prior to green up date & 48, 24 per phenotype
and 8 per phenotype \&~ genetic group \\
day length change (seconds) on a specific day indicated by the time
frame & 1, 2, 3, 4, 5, 6, 7, and 14 days prior to green up date & 48, 24
per phenotype and 8 per phenotype \& genetic group \\
average temperature in the time frame & 1, 2, 3, 4, 5, 6, 7, 14, 21, 28
days prior to green up date & 60, 30 per phenotype and 10 per phenotype
\& genetic group \\
\end{longtable}

Figures can be included using Quarto's syntax, code chunks, or with
LaTeX.

For simple figures, Quarto syntax is typically easiest. Full details are
available in \href{https://quarto.org/docs/authoring/figures.html}{the
Quarto documentation}.

\includegraphics{figs/frog.pdf}

\caption{\label{fig-frog}Placeholder image of a frog with a long example
legend to show justification setting.}

Tables can similarly be included using Quarto's syntax, code chunks, or
with LaTeX. I tend to use code chunks, but LaTeX works very well for
classic tables, like regression tables or summary statistics.

\begin{table}[t!]
\centering
\caption{Comparison of the fitted potential energy surfaces and ab initio benchmark electronic energy calculations}
\begin{tabular}{lrrr}
Species & CBS & CV & G3 \\
\midrule
1. Acetaldehyde & 0.0 & 0.0 & 0.0 \\
2. Vinyl alcohol & 9.1 & 9.6 & 13.5 \\
3. Hydroxyethylidene & 50.8 & 51.2 & 54.0\\
\bottomrule
\end{tabular}

\addtabletext{nomenclature for the TSs refers to the numbered species in the table.}
\end{table}

\subsubsection*{Digital Figures}\label{digital-figures}
\addcontentsline{toc}{subsubsection}{Digital Figures}

EPS, high-resolution PDF, and PowerPoint are preferred formats for
figures that will be used in the main manuscript. Authors may submit PRC
or U3D files for 3D images; these must be accompanied by 2D
representations in TIFF, EPS, or high-resolution PDF format. Color
images must be in RGB (red, green, blue) mode. Include the font files
for any text.

Images must be provided at final size, preferably 1 column width
(8.7cm). Figures wider than 1 column should be sized to 11.4cm or 17.8cm
wide. Numbers, letters, and symbols should be no smaller than 6 points
(2mm) and no larger than 12 points (6mm) after reduction and must be
consistent.

Figures and tables should be labelled and referenced in the standard way
using the \texttt{\textbackslash{}label\{\}} and
\texttt{\textbackslash{}ref\{\}} commands.

Figure Figure~\ref{fig-frog} shows an example of how to insert a
column-wide figure. To insert a figure wider than one column, please use
the \texttt{figure*} environment. Figures wider than one column should
be sized to 11.4 cm or 17.8 cm wide. Use \texttt{SCfigure*} environment
for a wide figure with side legends, such as Figure~\ref{fig-side}.

In Quarto, we can do these by setting the \texttt{fig-env} command to
\texttt{figure*} or \texttt{SCfigure*}

\begin{SCfigure*}

\includegraphics{figs/frog.pdf}

\caption{\label{fig-side}This legend would be placed at the side of the
figure, rather than below it.}

\end{SCfigure*}%

\subsection*{Tables}\label{tables}
\addcontentsline{toc}{subsection}{Tables}

Tables should be included in the main manuscript file and should not be
uploaded separately.

\subsection*{Single column equations}\label{single-column-equations}
\addcontentsline{toc}{subsection}{Single column equations}

Authors may use 1- or 2-column equations in their article, according to
their preference.

To allow an equation to span both columns, use the
\texttt{\textbackslash{}begin\{figure*\}...\textbackslash{}end\{figure*\}}
environment mentioned above for figures.

Note that the use of the \texttt{widetext} environment for equations is
not recommended, and should not be used.

\begin{figure*}[bt!]
\begin{align*}
(x+y)^3&=(x+y)(x+y)^2\\
       &=(x+y)(x^2+2xy+y^2) \numberthis \label{eqn:example} \\
       &=x^3+3x^2y+3xy^3+x^3.
\end{align*}
\end{figure*}

\section*{Material and Methods}\label{material-and-methods}
\addcontentsline{toc}{section}{Material and Methods}

\subsection*{Supporting Information Appendix
(SI)}\label{supporting-information-appendix-si}
\addcontentsline{toc}{subsection}{Supporting Information Appendix (SI)}

Authors should submit SI as a single separate SI Appendix PDF file,
combining all text, figures, tables, movie legends, and SI references.
SI will be published as provided by the authors; it will not be edited
or composed. Additional details can be found in the
\href{https://www.pnas.org/authors/submitting-your-manuscript\#manuscript-formatting-guidelines}{PNAS
Author Center}. The PNAS Overleaf SI template can be found
\href{https://www.overleaf.com/latex/templates/pnas-template-for-supplementary-information/wqfsfqwyjtsd}{here}.
Refer to the SI Appendix in the manuscript at an appropriate point in
the text. Number supporting figures and tables starting with S1, S2,
etc.

Authors who place detailed materials and methods in an SI Appendix must
provide sufficient detail in the main text methods to enable a reader to
follow the logic of the procedures and results and also must reference
the SI methods. If a paper is fundamentally a study of a new method or
technique, then the methods must be described completely in the main
text.

\subsubsection*{SI Datasets}\label{si-datasets}
\addcontentsline{toc}{subsubsection}{SI Datasets}

Supply .xlsx, .csv, .txt, .rtf, or .pdf files. This file type will be
published in raw format and will not be edited or composed.

\subsubsection*{SI Movies}\label{si-movies}
\addcontentsline{toc}{subsubsection}{SI Movies}

Supply Audio Video Interleave (avi), Quicktime (mov), Windows Media
(wmv), animated GIF (gif), or MPEG files. Movie legends should be
included in the SI Appendix file. All movies should be submitted at the
desired reproduction size and length. Movies should be no more than 10MB
in size.

\section{References}\label{references-1}

@@

\phantomsection\label{refs}
\begin{CSLReferences}{0}{1}
\end{CSLReferences}

>>>>

% TODO: Add custom LaTeX header directives here
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Switchgrass Phenology and Genotype-by-Environment Mapping},
  pdfauthor={Alice H. MacQueen; Li Zhang; Samuel A. Smith; Jason Bonnette; Arvid R. Boe; Phillip A. Fay; Felix B. Fritschi; David B. Lowry; Robert B. Mitchell; Francis M. Rouquette Jr; Yanqi Wu; Arbel Harpak; Thomas E. Juenger},
  pdfkeywords={allele-by-environment effect variation, antagonistic
pleiotropy, photoperiod, cumulative growing degree days, genetic
variation},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\title{Switchgrass Phenology and Genotype-by-Environment Mapping}

@@

\author[a%
,\equalcont%
,\correspond%
]{Alice H. MacQueen}
\author[a%
,\equalcont%
%
]{Li Zhang}
\author[a%
,\equalcont%
%
]{Samuel A. Smith}
\author[a%
%
%
]{Jason Bonnette}
\author[b%
%
%
]{Arvid R. Boe}
\author[c%
%
%
]{Phillip A. Fay}
\author[d%
%
%
]{Felix B. Fritschi}
\author[e%
%
%
]{David B. Lowry}
\author[f%
%
%
]{Robert B. Mitchell}
\author[g%
%
%
]{Francis M. Rouquette Jr}
\author[h%
%
%
]{Yanqi Wu}
\author[a%
%
%
]{Arbel Harpak}
\author[a%
%
,\correspond%
]{Thomas E. Juenger}

\affil[a]{University of Texas at Austin, Department of Integrative
Biology, Austin, 78712}
\affil[b]{South Dakota State University, Department of
Agronomy, Brookings}
\affil[c]{USDA-ARS, Grassland, Soil and Water Research
Laboratory, Temple}
\affil[d]{University of Missouri, Division of Plant Sciences, Columbia}
\affil[e]{Michigan State University, Department of Plant Biology, East
Lansing}
\affil[f]{USDA-ARS, Wheat, Sorghum, and Forage Research Unit, Lincoln}
\affil[g]{Texas A\&M University, Texas A\&M AgriLife Research and
Extension Center, Overton}
\affil[h]{Oklahoma State University, Department of Plant and Soil
Sciences, Stillwater}

\leadauthor{MacQueen}

@@

\authorcontributions{T.E.J. designed research. D.B.L. contributed plant
material and resources. J.B., D.B.L., and T.E.J. designed and executed
field experiments. A.R.B., P.A.F., F.B.F., D.B.L., R.B.M., F.M.R., Y.W.,
and T.E.J. hosted field experiments. A.H.M., L.Z., and S.A.S. conducted
statistical and computational analyses. The manuscript was written by
A.H.M. with contributions from all authors.}

\authordeclaration{The authors declare no conflicts of interest.}

\equalauthors{\textsuperscript{\equalcont} A.H.M. contributed equally to
this work with L.Z. and S.A.S.}

\correspondingauthor{\textsuperscript{\correspond}To whom correspondence should be addressed. E-mail: alicem@uchicago.edu}
\correspondingauthor{\textsuperscript{\correspond}To whom correspondence should be addressed. E-mail: tjuenger@utexas.edu}

@@

\begin{abstract}
The timing of vegetative growth and flowering in plants (``phenological
timings'') multiple depend on environmental cues; thus, phenological
traits can have important genotype-by-environment (GxE) interactions.
Here, we map genotype-by-weather effects on phenological timings in
multiple populations of two highly divergent switchgrass (Panicum
virgatum) populations spanning the central United States. We distinguish
GxWeather patterns covarying with interpretable weather-based cues and
agnostic, empirically-inferred GxWeather patterns. Most GxWeather
effects belong to the latter category and do not covary with
weather-based cues. However, \textgreater XX\% of Gulf population
effects on flowering time covary with photoperiod cues, while
\textgreater XX\% of Midwest upland population SNPs affecting flowering
had effects that covaried with XX cues. An independent pseudo-F2 cross
of Gulf and Midwest individuals mapped 23 additive QTLs for flowering at
the same common gardens, all with significant mash associations and ten
with enrichment of highly significant mash associations. We demonstrate
that we can identify QTL with GxE and assign them to specific
weather-based cues and/or data-driven patterns. Breeding for particular
alleles at these loci could change flowering responsiveness to
photoperiod cues in switchgrass. More broadly, this approach could be
used to identify genetic marker-environment interactions in any species
with related populations phenotyped in multiple environments.
\end{abstract}

@@

\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}
\firstpage[15]{2}

@@

Flowering time, in particular, is a common subject of GxE research
(5--11), a key output of selection driving adaptation to local
environments (2, 12, 13), and a selection target for crop improvement to
adapt crops to local or future environments (14). Changing flowering
responsiveness to photoperiod cues has allowed geographic range
expansion and increased yields in a number of cereal species (15--19)
and other crops (20, 21). Recent statistical advances in studying
phenological GxE have involved determining critical environmental
indices before the phenological event occurs, such as photothermal time
within a critical growth window (10). However, most studies of flowering
GxE focus on finding a single, best fitting form of genotype-environment
covariance, despite the key expectation that different genetic
subpopulations, and even different genomic regions, have likely evolved
distinct patterns of GxE. Additionally, despite theoretical predictions
that local adaptation should involve antagonistic pleiotropy, or
sign-changing GxE, at the level of individual loci (22--25), previous
work has found limited evidence of antagonistic pleiotropy (12, 26), but
has been limited by a known statistical bias that reduced detection of
SNP effects that differ in sign (26--28). Thus, despite substantial
interest in the frequencies of various forms of GxE, the prevalence of
antagonistic pleiotropy relative to other forms of GxE remains unknown.

Switchgrass (\emph{Panicum virgatum}) is considered a short-day plant
with reproductive development strongly linked to day of the year (29).
However, as part of its wide environmental adaptation across the eastern
half of North America, its photoperiodicity has been predicted to differ
by plant latitude of origin (30, 31). We previously found divergent
Midwest and Gulf genetic subpopulations of switchgrass which segregate
for distinct sets of environmental adaptations, based on the analysis of
two measures of fitness (i.e.,biomass and survival) among 732 genotypes
(32). The Midwest genetic subpopulation is primarily composed of
individuals from the well-studied upland switchgrass ecotype (33, 34),
while the Gulf subpopulation has individuals from the lowland ecotype
and the phenotypically intermediate coastal ecotype (32).

Here, we test how the Midwest and Gulf subpopulations differ in their
phenological adaptations across environments, and hence their
phenological GxE. We phenotype a diversity panel of hundreds of
switchgrass genotypes from the Midwest and Gulf subpopulations for the
timing of vegetative development (``green-up'') and reproductive
development (flowering) at eight common garden locations spanning 17
degrees of latitude. These gardens cover the majority of the latitudinal
and climatic range of switchgrass and capture the most comprehensive
picture to date of the environmental variation this species encounters.
After single-site estimation of SNP effects for these phenotypes, we use
multivariate adaptive shrinkage (mash) to jointly estimate SNP effects
in each subpopulation and phenotype at all eight sites (35). Mash allows
us to identify and specify multiple ways genetic effects linked to SNPs
may covary with the environment, and does not have a statistical bias in
detecting SNP effects with the same or opposite signs (36). To confirm
our genetic mapping of GxE, we compare to mapping results from an
outbred pseudo-F2 cross grown at the same sites. Taken together, our
results allow us to describe the environmental cues and genetic
variation affecting phenology in two divergent natural populations of
switchgrass.

@@

In our diversity panel of tetraploid switchgrass (32), genotypes from
the Gulf and Midwest genetic subpopulations had distinct phenological
timings and distinct patterns of phenological correlations across our
eight common garden sites (Figure~\ref{fig-map}). At the three Texas
common gardens (hereafter `Texas' gardens), located within the natural
range of the Gulf subpopulation, Gulf green-up occurred before
Midwestern green-up, and Gulf flowering occurred after Midwestern
flowering (Fig. 1A). At the four northernmost common gardens (hereafter
`North' gardens), located within the natural range of the Midwest
subpopulation, both Gulf green-up and flowering occurred after Midwest
green-up and flowering. At the Oklahoma common garden, located near the
natural range limits of both the Gulf and the Midwest subpopulations,
Gulf and Midwest green-up occurred over the same time period. These
patterns led to strong negative phenotypic correlations for green-up
between the North and Texas gardens, particularly in the Gulf and Both
subpopulations, and contributed to positive phenotypic correlations for
flowering time of larger magnitude at more northern gardens (Fig. 1B).

@@

\begin{table}[t!]
\centering
\caption{Weather variables and time frames around the green up and flowering date used to construct the hypothesis-based covariance matrices. The correlations between values of these weather variables for genetically identical plants grown in different gardens were used to fill off-diagonal cells of the covariance matrices. Narrow-sense heritabilities for these values at each garden were used for the diagonal cells.}
\begin{tabular}{lrr}
Weather variable & Time Frame & Covariance Matrices tested \\
\midrule
1. cumulative GDD at 12C in the time frame & (1-7), 14, 21, 28 days prior to green up date & 60 \\
2. cumulative rainfall in the time frame & (1-7), 14, 21, 28 days prior to green up date & 60  \\
3. day length (hours) on a specific day indicated by the time frame & (1-7) and 14 days prior to green up date & 48 \\
4. day length change (seconds) on a specific day indicated by the time frame & (1-7) and 14 days prior to green up date & 48 \\
5. average temperature in the time frame & (1-7), 14, 21, 28 days prior to green up date & 60 \\
\bottomrule
\end{tabular}

\end{table}

\includegraphics{images/Fig1_200dpi.png}

\caption{\label{fig-map}Figure 1. Characterization of green-up and
flowering dates from the switchgrass diversity panel. (a) Map and trait
histograms of green-up and flowering dates across two genetically
distinct switchgrass subpopulations and eight common gardens. Purple
represents individuals from the Midwest genetic subpopulation, and pink
individuals from the Gulf subpopulation. Vertical dashed lines indicate
the summer solstice. Common gardens are arranged in latitudinal order.
(b) Phenotypic correlations between clonal replicates planted at eight
common gardens, within and between two genetic subpopulations. (c)
Narrow sense heritability of green-up and flowering within single common
gardens (purple) and across all eight common gardens (green), within and
between two genetic subpopulations. (d) Flow diagram of the methods
applied to the green-up and flowering dates to jointly estimate SNP
effects across all sites. Mash was fit to SNP effect data and used to
find covariance matrices that improved the mash model likelihood using a
large set of randomly selected, relatively unlinked SNP effects; this
model was applied to a ``strong'' set of SNP effects with large effect
sizes in the univariate GWAS.}

\includegraphics{images/Figure_2_Covariance_matrices_and_posterior_weights.png}

\caption{\label{fig-covar}Figure 2.}

@@

\section{References}\label{references}

@@

\phantomsection\label{refs}
\begin{CSLReferences}{0}{1}
\bibitem[\citeproctext]{ref-bauerle_photoperiodic_2012}
\CSLLeftMargin{1. }%
\CSLRightInline{W. L. Bauerle, \emph{et al.},
\href{https://doi.org/10.1073/pnas.1119131109}{Photoperiodic regulation
of the seasonal pattern of photosynthetic capacity and the implications
for carbon cycling}. \emph{Proceedings of the National Academy of
Sciences} \textbf{109}, 8612--8617 (2012).}

\bibitem[\citeproctext]{ref-andres2012genetic}
\CSLLeftMargin{2. }%
\CSLRightInline{F. Andrés, G. Coupland, The genetic basis of flowering
responses to seasonal cues. \emph{Nature Reviews Genetics} \textbf{13},
627--639 (2012).}

\bibitem[\citeproctext]{ref-korner2010phenology}
\CSLLeftMargin{3. }%
\CSLRightInline{C. Körner, D. Basler, Phenology under global warming.
\emph{Science} \textbf{327}, 1461--1462 (2010).}

\bibitem[\citeproctext]{ref-botero2015evolutionary}
\CSLLeftMargin{4. }%
\CSLRightInline{C. A. Botero, F. J. Weissing, J. Wright, D. R.
Rubenstein, Evolutionary tipping points in the capacity to adapt to
environmental change. \emph{Proceedings of the National Academy of
Sciences} \textbf{112}, 184--189 (2015).}

\bibitem[\citeproctext]{ref-blackman2013interacting}
\CSLLeftMargin{5. }%
\CSLRightInline{B. K. Blackman, Interacting duplications, fluctuating
selection, and convergence: The complex dynamics of flowering time
evolution during sunflower domestication. \emph{Journal of experimental
botany} \textbf{64}, 421--431 (2013).}

\bibitem[\citeproctext]{ref-henry2014transitions}
\CSLLeftMargin{6. }%
\CSLRightInline{L. P. Henry, R. H. Watson, B. K. Blackman, Transitions
in photoperiodic flowering are common and involve few loci in wild
sunflowers (helianthus; asteraceae). \emph{American Journal of Botany}
\textbf{101}, 1748--1758 (2014).}

\bibitem[\citeproctext]{ref-aagren2017adaptive}
\CSLLeftMargin{7. }%
\CSLRightInline{J. Ågren, C. G. Oakley, S. Lundemo, D. W. Schemske,
Adaptive divergence in flowering time among natural populations of
arabidopsis thaliana: Estimates of selection and QTL mapping.
\emph{Evolution} \textbf{71}, 550--564 (2017).}

\bibitem[\citeproctext]{ref-brachi2010linkage}
\CSLLeftMargin{8. }%
\CSLRightInline{B. Brachi, \emph{et al.}, Linkage and association
mapping of arabidopsis thaliana flowering time in nature. \emph{PLoS
genetics} \textbf{6}, e1000940 (2010).}

\bibitem[\citeproctext]{ref-dittmar2014flowering}
\CSLLeftMargin{9. }%
\CSLRightInline{E. L. Dittmar, C. G. Oakley, J. Ågren, D. W. Schemske,
Flowering time QTL in natural populations of arabidopsis thaliana and
implications for their adaptive value. \emph{Molecular ecology}
\textbf{23}, 4291--4303 (2014).}

\bibitem[\citeproctext]{ref-li2018genomic}
\CSLLeftMargin{10. }%
\CSLRightInline{X. Li, T. Guo, Q. Mu, X. Li, J. Yu, Genomic and
environmental determinants and their interplay underlying phenotypic
plasticity. \emph{Proceedings of the National Academy of Sciences}
\textbf{115}, 6679--6684 (2018).}
